Map{R}educe in {MPI} for Large-scale Graph Algorithms

### 2010-11-19 paper/mr.tex
git-svn-id: svn+ssh://software-srn.sandia.gov/svn/mapreduce/trunk@672 3fc257bb-8550-0410-81b3-a5c103da83f7

The basic datums stored and operated on by any MapReduce framweork are
key/value (KV) pairs.  In the MR-MPI library, individual keys or
values can be of any data type or length, or combinations of multiple
types (one integer, a string of characters, two integers and a double,
etc); they are simply treated as byte strings by the library.  A KV
pair always has a key; its value may be NULL.  KV pairs are stored
within a MapReduce (MR) object.  A user program may create one or more
MR objects to implement an algorithm.  Various MapReduce operations
(map, reduce, etc) are invoked on an object and its KV pairs; KV pairs
can also be passed between and combined across objects.

A related data type in the MR-MPI library is the key/multivalue (KMV)
pair, where all values associated with the same key are collected and
stored contiguously as a multivalue, which is just a longer byte
string with an associated vector of lengths, one integer length per
value.  In this section, we assume the KV and KVM datums operated on
by each processor fit in local memory.

@@

The interface to various low-level operations is provided so that a
user's program can string them together in various ways to construct
usefule MapReduce algorithms.  For example, output from a {\it
reduce()} can serve as input to a subsequent {\it map()} or {\it
collate()}.  Likewise, data can be partitioned across multiple
MapReduce (MR) objects, each with their own set of KV pairs.  For
example, as will be illustrated in Section ~\ref{sec:graph}, one MR
object can store the edges of a graph, and another its vertices.  KV
pairs from multiple MR objects can be combined to perform new
sequences of {\it map()}, {\it collate()}, and {\it reduce()}
operations.

>>>>

The basic datums stored and operated on by any MapReduce framework are
key/value (KV) pairs.  In the MR-MPI library, individual keys or
values can be of any data type or length, or combinations of multiple
types (one integer, a string of characters, two integers and a double,
etc); they are simply treated as byte strings by the library.  A KV
pair always has a key; its value may be NULL.  KV pairs are stored
within a MapReduce (MR) object.  A user program may create one or more
MR objects to implement an algorithm.  Various MapReduce operations
(map, reduce, etc) are invoked on an object and its KV pairs; KV pairs
can also be passed between and combined across objects.

A related data type in the MR-MPI library is the key/multivalue (KMV)
pair, where all values associated with the same key are collected and
stored contiguously as a multivalue, which is just a longer byte
string with an associated vector of lengths, one integer length per
value.  In this section, we assume the KV and KMV datums operated on
by each processor fit in local memory.

@@

The interface to various low-level operations is provided so that a
user's program can string them together in various ways to construct
useful MapReduce algorithms.  For example, output from a {\it
reduce()} can serve as input to a subsequent {\it map()} or {\it
collate()}.  Likewise, data can be partitioned across multiple
MapReduce (MR) objects, each with their own set of KV pairs.  For
example, as will be illustrated in Section ~\ref{sec:graph}, one MR
object can store the edges of a graph, and another its vertices.  KV
pairs from multiple MR objects can be combined to perform new
sequences of {\it map()}, {\it collate()}, and {\it reduce()}
operations.

### 2010-11-19 paper/paper.tex
git-svn-id: svn+ssh://software-srn.sandia.gov/svn/mapreduce/trunk@672 3fc257bb-8550-0410-81b3-a5c103da83f7

We describe a parallel library written with message-passing (MPI)
calls that allows algorithms to be expressed in the MapReduce
paradigm.  This means the calling program does not need to include
explicit parallel code, but instead provides ``map'' and ``reduce''
functions that operate independently on elements of a data set
distributed across processors.  The library performs needed data
movement between processors.  We describe how typical MapReduce
funtionality can be implemented in an MPI context, and also in an
out-of-core manner for data sets that do not fit within the aggregate
memory of a parallel machine.  Our motivation for creating this
library was to enable graph algorithms to be written as MapReduce
operations, allowing processing of terabyte-scale data sets.  We
outline MapReduce versions of several such algorithms: vertex ranking
via PageRank, triangle finding, connected component identification,
Luby's algorithm for maximally independent sets, and single-source
shortest-path calculation.  To test the algorithms on arbitrarily
large artificial graphs we generate randomized R-MAT matrices in
parallel; a MapReduce version of this operation is also described.
Performance and scalability results for the various algorithms are
presented for varying size graphs on a distributed-memory cluster.
For some cases, we compare the results with non-MapReduce algorithms,
different machines, and different MapReduce software, namely Hadoop.
Our open-source library is written in C++, is callable from C++, C,
Fortran, or scripting languages such as Python, and can run on any
parallel platform that supports MPI.

>>>>

We describe a parallel library written with message-passing (MPI)
calls that allows algorithms to be expressed in the MapReduce
paradigm.  This means the calling program does not need to include
explicit parallel code, but instead provides ``map'' and ``reduce''
functions that operate independently on elements of a data set
distributed across processors.  The library performs needed data
movement between processors.  We describe how typical MapReduce
functionality can be implemented in an MPI context, and also in an
out-of-core manner for data sets that do not fit within the aggregate
memory of a parallel machine.  Our motivation for creating this
library was to enable graph algorithms to be written as MapReduce
operations, allowing processing of terabyte-scale data sets.  We
outline MapReduce versions of several such algorithms: vertex ranking
via PageRank, triangle finding, connected component identification,
Luby's algorithm for maximally independent sets, and single-source
shortest-path calculation.  To test the algorithms on arbitrarily
large artificial graphs we generate randomized R-MAT matrices in
parallel; a MapReduce version of this operation is also described.
Performance and scalability results for the various algorithms are
presented for varying size graphs on a distributed-memory cluster.
For some cases, we compare the results with non-MapReduce algorithms,
different machines, and different MapReduce software, namely Hadoop.
Our open-source library is written in C++, is callable from C++, C,
Fortran, or scripting languages such as Python, and can run on any
parallel platform that supports MPI.

### 2010-11-19 paper/results.tex
git-svn-id: svn+ssh://software-srn.sandia.gov/svn/mapreduce/trunk@672 3fc257bb-8550-0410-81b3-a5c103da83f7

In this section, we present performance results for the graph
algorithms of the preceeding section, implemented as small C++
programs calling our MR-MPI library.  The benchmarks were run on a
medium-sized Linux cluster of 2 GHz dual-core AMD Opteron processors
connected via a Myrinet network.  Most importantly for MR-MPI, each
node of the cluster has one local disk, which is used for out-of-core
operations.  To avoid contention for disk I/O, we ran all experiments
with one MPI process per node.  For comparisons with other
implementations, we used either the same cluster or, where noted,
Sandia's Cray XMT, a multi-threaded parallel computer with 500 {MHz}
processors and a 3D-Torus network.

@@

All data sets used R-MAT parameters $(a, b, c, d) = (0.57, 0.19, 0.19,
0.05)$ and generated 8 edges per vertex (on average).  These
parameters create a highly skewed degree distribution, as indicated by
the maximumn vertex degree in Table~\ref{t:rmats}.

@@

In Figure~\ref{f:cc}, we show the performance of the connected
component identification algorithm (Figure~\ref{fig:cc}).  A
comparison is made with a hybrid ``Giant Connected Component''
implementation using both Trilinos and MR-MPI.  Power law graphs often
have one or two very large components and many very small components.
The hybrid algorithm exploits this feature by using an inexpensive
breadth-first search (BFS) from the vertex with highest degree to
identify the largest components one at a time, followed by a more
expensive algorithm to identify the small components in the remainder
of the graph.  In our hybrid implementation, we intially perform
matrix-vector multiplications in Trilinos to perform a BFS, finding a
handful of components that include 70\% or more of the vertices (for
these R-MAT matrices).  We then apply our MapReduce algorithm to the
remaining graph to identify the small components.  This hybrid
approach is quite effective, reducing the execution time to identify
all components in RMAT-20 and RMAT-24 by 80-99\%, as shown in
Figure~\ref{f:cc}.  For RMAT-28, where the graph is too large to fit
into memory, Trilinos cannot perform the initial BFS, so only the
MapReduce algorithm was used.  The total number of components
identified were roughly 498K, 9.37M, and 169M for the three increasing 
R-MAT sizes; five or six iterations were 
required by the algorithm.

@@

\begin{figure}[htb]
\includegraphics[width=\textwidth]{fig_ssspA.pdf}
\caption{SSSP using MR-MPI for WebGraphA with
13.2M vertices and 31.9M edges.  Runtimes using Hadoop and PBGL
are also shown.}
\label{f:ssspA}
\end{figure}

@@

For comparison, we wrote a parallel distributed-memory R-MAT
generation code with explicit MPI calls using a different
(non-MapReduce) algorithm than that of Figures \ref{fig:rmat} and
\ref{fig:rmat2}.  It required 1006 lines of code, mainly because it
must implement the parallelism explicitly whereas the codes of this
section relegate that task to the MR-MPI library.  A similar all-MPI
implementation of the PageRank algorithm required 1206 lines.  By
contrast an implementation of PageRank using the Trilinos parallel
linear algebra library (discussed in Section
\ref{subsec:results_pagerank}) required only 153 lines of code, since
an entire matrix-vector multiply can be invoked by a single call to
Trilinos.  Our collaborators (see Acknowledgements) have a Hadoop
implementation of the SSSP algorithm of Figure \ref{fig:sssp}.  It
required XXX lines of Java code.

>>>>

In this section, we present performance results for the graph
algorithms of the preceding section, implemented as small C++
programs calling our MR-MPI library.  The benchmarks were run on a
medium-sized Linux cluster of 2 GHz dual-core AMD Opteron processors
connected via a Myrinet network.  Most importantly for MR-MPI, each
node of the cluster has one local disk, which is used for out-of-core
operations.  To avoid contention for disk I/O, we ran all experiments
with one MPI process per node.  For comparisons with other
implementations, we used either the same cluster or, where noted,
Sandia's Cray XMT, a multi-threaded parallel computer with 500 {MHz}
processors and a 3D-Torus network.

@@

All data sets used R-MAT parameters $(a, b, c, d) = (0.57, 0.19, 0.19,
0.05)$ and generated 8 edges per vertex (on average).  These
parameters create a highly skewed degree distribution, as indicated by
the maximum vertex degree in Table~\ref{t:rmats}.

@@

In Figure~\ref{f:cc}, we show the performance of the connected
component identification algorithm (Figure~\ref{fig:cc}).  A
comparison is made with a hybrid ``Giant Connected Component''
implementation using both Trilinos and MR-MPI.  Power law graphs often
have one or two very large components and many very small components.
The hybrid algorithm exploits this feature by using an inexpensive
breadth-first search (BFS) from the vertex with highest degree to
identify the largest components one at a time, followed by a more
expensive algorithm to identify the small components in the remainder
of the graph.  In our hybrid implementation, we initially perform
matrix-vector multiplications in Trilinos to perform a BFS, finding a
handful of components that include 70\% or more of the vertices (for
these R-MAT matrices).  We then apply our MapReduce algorithm to the
remaining graph to identify the small components.  This hybrid
approach is quite effective, reducing the execution time to identify
all components in RMAT-20 and RMAT-24 by 80-99\%, as shown in
Figure~\ref{f:cc}.  For RMAT-28, where the graph is too large to fit
into memory, Trilinos cannot perform the initial BFS, so only the
MapReduce algorithm was used.  The total number of components
identified were roughly 498K, 9.37M, and 169M for the three increasing
R-MAT sizes; five or six iterations were required by the algorithm.

@@

\begin{figure}[htb]
\includegraphics[width=\textwidth]{fig_ssspA.pdf}
\caption{SSSP using MR-MPI for WebGraphA with 13.2M vertices and 31.9M
  edges.  Run times using Hadoop and PBGL are also shown.}
\label{f:ssspA}
\end{figure}

@@

For comparison, we wrote a parallel distributed-memory R-MAT
generation code with explicit MPI calls using a different
(non-MapReduce) algorithm than that of Figures \ref{fig:rmat} and
\ref{fig:rmat2}.  It required 1006 lines of code, mainly because it
must implement the parallelism explicitly whereas the codes of this
section relegate that task to the MR-MPI library.  A similar all-MPI
implementation of the PageRank algorithm required 1206 lines.  By
contrast an implementation of PageRank using the Trilinos parallel
linear algebra library (discussed in Section
\ref{subsec:results_pagerank}) required only 153 lines of code, since
an entire matrix-vector multiply can be invoked by a single call to
Trilinos.  Our collaborators (see Acknowledgments) have a Hadoop
implementation of the SSSP algorithm of Figure \ref{fig:sssp}.  It
required XXX lines of Java code.
